---
title: "SWIFT spatial: descriptive analysis of predictor variables"
author: "Christine Tedijanto christine.tedijanto@ucsf.edu"
date: "Last updated: 09/21/2021, cleaned: 02/28/2022"
output:
    html_document:
      toc: true
      toc_float: true
      toc_collapsed: true
      toc_depth: 3
      code_folding: hide
      df_print: paged
number_sections: false
---

**Objective:** Combine WUHA measurements and geospatial features into dataset for analysis.

```{r setup, message = FALSE}
library(here)
source(here("0-config.R"))

# read in data
ind_all <- read_csv(here(data_path, "ind_all.csv"))
clu_random_byage <- read_csv(here(data_path, "clu_random_byage.csv"))
swift_spatial_features_annual <- readRDS(here(data_path, "swift_spatial_features_annual.rds"))
swift_spatial_features_monthly <- readRDS(here(data_path, "swift_spatial_features_monthly.rds"))
swift_spatial_features_static <- readRDS(here(data_path, "swift_spatial_features_static.rds"))
load(here(data_path, "swift_spatial_data.Rda"))

# list of months (yyyy-mm) from 2015 through 2018
start_date_months <- start_dates %>% substr(start = 1, stop = 7)
```

### Helper functions

```{r centroids}
### pull values of spatial features by cluster centroids -----
# features_sf: simple features dataset with features and corresponding geometries
# centroids: simple features dataset with cluster ids and cluster centroids
# returns data frame with cluster ids and corresponding feature values
get_centroid_values <- function(features_sf, centroids){
  
  # check if both features and centroids are sf
  if(("sf" %in% class(features_sf)) & ("sf" %in% class(centroids))){
    
    ret <- centroids %>%
      st_join(features_sf, join = st_intersects, left = TRUE) %>%
      st_drop_geometry()
    
    return(ret)
  
  } else {
    print("both arguments must be spatial features objects")
  }

}
```

### Explore time-varying features

Here, we compare different summaries of time-varying variables. If highly correlated, we will use just a single summary (e.g. annual) of the variable in our predictive model.

```{r time_vars, fig.width = 8.5, fig.height = 10}
### create summary (mean) vector of monthly values -----
# df: dataframe where columns are stored, columns should be named <var>_<month>
# var: variable to summarize over
# months: months to summarize over
# return vector of summary (same order as df rows)
get_time_summary <- function(df, var, months){
  
  if("sf" %in% class(df)){
    df <- df %>% st_drop_geometry()
  }
  
  if(length(months) == 1){
    ret <- df[,paste0(var,"_",months)]
  } else {
    ret <- rowMeans(df[,paste0(var,"_",months)], na.rm = TRUE)
  }
  
  return(ret)
  
}

plot_cor_over_time <- function(var){
  
  # get summaries over last few months for given variable
  # apply over each survey time point
  monthly_df <- lapply(survey_list,
                       function(x){
                         swift_spatial_features_monthly %>%
                           st_drop_geometry() %>%
                           mutate(last1mo = get_time_summary(., var, start_date_months[c(12)+x]),
                                  last2mo = get_time_summary(., var, start_date_months[c(11:12)+x]),
                                  last3mo = get_time_summary(., var, start_date_months[c(10:12)+x]),
                                  last4mo = get_time_summary(., var, start_date_months[c(9:12)+x]),
                                  last6mo = get_time_summary(., var, start_date_months[c(7:12)+x]),
                                  last8mo = get_time_summary(., var, start_date_months[c(5:12)+x]),
                                  last10mo = get_time_summary(., var, start_date_months[c(3:12)+x]),
                                  last12mo = get_time_summary(., var, start_date_months[c(1:12)+x])) %>% 
                           dplyr::select(starts_with("last"))
                         }) %>% 
    bind_rows()

  # plot
  monthly_df %>% 
    dplyr::select(starts_with("last")) %>% 
    cor(method = "pearson", use = "complete.obs") %>%
    corrplot.mixed(lower = "number",
                   upper = "ellipse",
                   tl.pos = "d",
                   tl.col = "black",
                   number.cex = 0.7,
                   tl.cex = 0.7,
                   title = var,
                   mar = c(0,0,2,0))
}

par(mfrow = c(3,2))
plot_cor_over_time("surface_water") # surface water
plot_cor_over_time("precip") # precipitation
plot_cor_over_time("tmax") # max temp
plot_cor_over_time("evi") # evi 
plot_cor_over_time("avg_rad") # night light radiance
```

Based on the correlation plots, we will include the following summaries:

-   surface_water: last 1 and 12 months
-   precip: last 1, 2, 4, 12 months
-   tmax: last 12 months
-   evi: last 12 months
-   avg_rad: last 12 months

### Create merged dataset

```{r merge, fig.height = 8.5, fig.width = 8.5}
## create variable time summaries listed above -----
temp_clu_features <- lapply(
  survey_list,
  function(x){
    swift_spatial_features_monthly %>%
      transmute(surface_water_1m = get_time_summary(., "surface_water", start_date_months[c(12)+x]),
                surface_water_12m = get_time_summary(., "surface_water", start_date_months[c(1:12)+x]),
                precip_1m= get_time_summary(., "precip", start_date_months[c(12)+x]),
                precip_2m = get_time_summary(., "precip", start_date_months[c(11:12)+x]),
                precip_4m = get_time_summary(., "precip", start_date_months[c(9:12)+x]),
                precip_12m = get_time_summary(., "precip", start_date_months[c(1:12)+x]),
                tmax_12m = get_time_summary(., "tmax", start_date_months[c(1:12)+x]),
                evi_12m = get_time_summary(., "evi", start_date_months[c(1:12)+x]),
                avg_rad_12m = get_time_summary(., "avg_rad", start_date_months[c(1:12)+x])) %>% 
      mutate(survey = x)}) %>%
  bind_rows() %>% 
  st_join(swift_spatial_features_static, join = st_equals_exact, par = 0.00001)

clu_features <- get_centroid_values(features_sf = temp_clu_features, centroids = cluster_sf)
```

```{r}
## merge in outcomes and coordinates to create full dataset -----
clu_random_modeldata <- clu_random_byage %>%
  left_join(clu_features, by = c("cluster_id", "survey")) %>% 
  left_join(cluster_sf %>%
              rowwise() %>% 
              mutate(lon = geometry[[1]][1],
                     lat = geometry[[1]][2]) %>% 
              ungroup() %>% 
              st_drop_geometry(),
            by = "cluster_id")
```

### Plot variable distributions

```{r, fig.width = 8.5, fig.height = 11}
clu_random_modeldata %>% 
  filter(age_group == "0-5y") %>%
  dplyr::select(-c(cluster_id, cluster_id_public, survey,
                   age_group, arm, source_pcr, starts_with("n_"))) %>%
  pivot_longer(cols = everything(), names_to = "col_var", values_to = "value") %>%
  # order facets for readability
  mutate(col_var = factor(col_var, levels = c("prevalence_pcr",
                                              "prevalence_clin",
                                              "prevalence_sero",
                                              "prevalence_Pgp3",
                                              "prevalence_Ct694",
                                              "prevalence_tf",
                                              "prevalence_ti",
                                              "precip_1m",
                                              "precip_2m",
                                              "precip_4m",
                                              "precip_12m",
                                              "tmax_12m",
                                              "surface_water_1m",
                                              "surface_water_12m",
                                              "evi_12m",
                                              "elevation_2000",
                                              "slope_2000",
                                              "osm_residential_area_2020",
                                              "hdx_building_count_2020",
                                              "pop_2018",
                                              "women_2019",
                                              "under_5_2019",
                                              "dist_to_road_2020",
                                              "accesshc_2019",
                                              "accesshcwalk_2019",
                                              "avg_rad_12m"))) %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_wrap(.~col_var, scales = "free", ncol = 4) +
  theme_minimal()
```

### Save dataset

```{r save}
saveRDS(object = clu_random_modeldata,
        file = here(data_path, "clu_random_modeldata.rds"))
```

### Session info

```{r}
sessionInfo()
```
